Collocated Deduplication with Fault Isolation for Virtual Machine Snapshot Backup

### 2013-07-05 snapshot_system/challenge.tex
added srb comparison text, updated references throughout paper.
chagne

As discussed earlier, collocating the backup service on the existing
cloud cluster avoids the extra cost to acquire a dedicated backup facility
and reduces the network bandwidth consumption in transferring the un-deduplicated
raw data for backup. 
Figure~\ref{fig:collocated} illustrates the cluster architeture where
each physical runs a backup service and a distribued file system (DFS)~\cite{GFS2004,Hadoop} 
serves a backup store  for the snapshopts.
The previous study shows that 
 deduplication can compress the backup copies 
effectively in a 10:1 or even 15:1 rage. 
Therefore  the portion of space in a cluster
allocated for snapshots of data should not dominate
the cluster storage usage.

@@

\begin{figure}[htb]
    \centering
    \includegraphics[width=3in]{images/colocated-arch.pdf}
    \label{fig:collocated}
    \caption{Collocated VM Backup System.}
\end{figure}

@@

We discuss the design considerations as follows. 
\begin{itemize}
\item {\em Deduplication localization, sharing minmization, and fault tolerance.}

%Previous work in deduplication~\cite{extreme_binning09,sparseindex09} has
%studied the solutions to reduce memory requirements.
%One important issue not considered in the previous work is how 
Because a data chunk is compared with signatures collected from other VMs during
the deduplication process, only one copy of duplicates is stored in the backup storage
and this artificially creates data dependency among different VM users. 
Content sharing via deduplication affects fault isolation since machine failures happen at daily 
basis in a large-scale cloud and
%failure of a shared block affects many VMs.
%the loss of a shared chunks can affect many VM users whose snapshots share this 
%data block. 
loss of a small number of shared data chunks can 
cause the unavailability of snapshots for a large number of virtual machines.
Localizing the impact of deduplication can incrase fault isolation and resilience.
Thus from the fault tolerance point of view,  duplicate sharing among mutiple VMs is 
discourage. On the other hand, we need to seek for a tradeoff since there are
a significant number of duplicates cross VMs.

% since storage for shared identical content chunks can become a failure point.
%Without any control of such data sharing, we can only increase  
%replication for global datasets to enhance the availability,
%but this incurs significantly higher costs.
%Thus a key goal  in our design
%is restricting the impact of  data loss to a limited number of VMs. 
%It is not desirable that small scale of data failure affects the backup of many VMs.
%when users access snapshots in a recovery process. 
\item{\bf Packaging  data chunks as file system blocks.}
Since the file block size in the Hadoop and GFS is uniform and large with 64MB as a default setting,
the content chunk in a deduplication system is of nonuniform size with 4KB or 8KB on average.
We need to build an intermediate layer that supports large snapshot writing with append operations
and infrequent snapshot access.
Packaging that maps data chunks to file system blocks can create data dependence among VMs
since a file block can be shared even more VMs.
Thus we need to consider a minimum assoocation of a fie system block to VMs in the packaging process.
%The distributed file system~\cite{GoogleFS,Hadoop} provides the replication support
%and we can leverage existing open-source software to implement such a layer.
%When the number of machines failed exceeds the replication degree, there could be
% some data loss.  
%Storage deduplication worsens the impact of such faults
%because  the majority of each snapshot (e.g. over 90\%) is deduplicated
%and the corresponding data chunks stored in the underlying file blocks
%are shared by many VMs. 
%The failure of one data chunk or block can impact many sharing VMs. 
\end{itemize}

@@

A deduplication scheme compares the fingerprints of the current snapshot
with its parent snapshot and also other snapshots in the entire cloud without consideration of .
%The traditional approach compares all fingerprints and
%stores one copy for all of a chunk's duplicates.
We call this as  the VM-oblivious (VO) approach.
In designing and selecting a duplication algorithm, we have considered the following options.
\begin{itemize}
\item {\bf Version-based change detection}.
VM snapshots can be  backed up  incrementally by identifying file  blocks that have
changed from the previous version of the snapshot~\cite{Clements2009,Vrable2009,TanIPDPS2011}.
Active snapshots will contain all the information needed to restore the virtual disk
and when deleting a snapshot, data referenced by other snapshots are removed.
While full signature comparison can deliver additional reduction~\cite{Guo2011,Dong,ExtremeBining},
content change detection falls into a VM-centric approach since deduplication is localizable. 
We are seeking for additional optimzations to improve duplication efficency and design the 
assoication of data chunks with the underlying file blocks so that
file block sharing among VMs is minimized.
\item {\bf sampled Index}.
One alternative approach to reducing the use of memory space is 
to use a sampled index with prefetching, proposed  by Guo and Efstathopoulos\cite{Guo2011}. 
An evaluation with our test data shows that using a sampled index can achieve a high
deduplication effiency with a single machine setup.  
One key problem is that the algorithm is VM oblivious and not easy to be adopted for VM-centric.
Another problem is  to design a distributed version 
in deduplicating large bodies of data. 
%The algorithm stores the sampled index in memory, which is required for high
%throughput to avoid the disk-bottleneck problem - since the index itself has no
%locality information, lookups are essentially random, so the index must be
%somehow optimized to prevent excessive disk seeking (the sampling algorithm does
%this by storing the index in memory). 
To use a distributed memory version of the sampled index, every deduplicate request
may access a remote machine for index lookup,  which  incurs a significant overhead.

Another possibiliy is to use this approach partially  in our VM centric solution by indexing
the most popular data chunks. For a small set of popular data chunks, the prefetching strategy
used in the sampled index will not work well because the spatial locality is limited among popular 
data chunks. In our test data,  on average the number of consecutive data chunks is 7, which is too small
to be effective for index sampling.

@@

With these considerations in mind, we study a 
VM-centric appraoch (called VC)
for a backup service co-hosted   in the existing set of machines and resource usage is friendly
to the existing applications.  
We will first discuss and analyze the integration of the VM-centric deduplication strategies with fault isolation, and then present
an architecture and implementation design with deletion support.

>>>>

As discussed earlier, collocating the backup service on the existing
cloud cluster avoids the extra cost to acquire a dedicated backup facility
and reduces the network bandwidth consumption in transferring the un-deduplicated
raw data for backup. 
Figure~\ref{fig:collocated} illustrates the cluster architecture where
each physical runs a backup service and a distributed file system (DFS)~\cite{GFS2004,Hadoop} 
serves a backup store  for the snapshots.
The previous study shows that 
 deduplication can compress the backup copies 
effectively in a 10:1 or even 15:1 rage. 
Therefore  the portion of space in a cluster
allocated for snapshots of data should not dominate
the cluster storage usage.

@@

\begin{figure}[htb]
    \centering
    \includegraphics[width=3in]{images/colocated-arch.pdf}
    \caption{Collocated VM Backup System.}
    \label{fig:collocated}
\end{figure}

@@

We discuss the design considerations as follows. 
\begin{itemize}
\item {\em Deduplication localization, sharing minimization, and fault tolerance.}

%Previous work in deduplication~\cite{extreme_binning09,sparseindex09} has
%studied the solutions to reduce memory requirements.
%One important issue not considered in the previous work is how 
Because a data chunk is compared with signatures collected from other VMs during
the deduplication process, only one copy of duplicates is stored in the backup storage
and this artificially creates data dependency among different VM users. 
Content sharing via deduplication affects fault isolation since machine failures happen at daily 
basis in a large-scale cloud and
%failure of a shared block affects many VMs.
%the loss of a shared chunks can affect many VM users whose snapshots share this 
%data block. 
loss of a small number of shared data chunks can 
cause the unavailability of snapshots for a large number of virtual machines.
Localizing the impact of deduplication can increase fault isolation and resilience.
Thus from the fault tolerance point of view,  duplicate sharing among multiple VMs is 
discourage. 
Another disadvantage of sharing is that it complicates snapshot deletion.
The mark-sweep approach~\cite{mark-sweep} is effective for deletion, and its main cost
is to count if a data chunk is still shared by other snapshots.
Thus our goal is to localize deduplication and minimize sharing, while we seek for 
a tradeoff since there are a significant number of duplicates cross VMs.

% since storage for shared identical content chunks can become a failure point.
%Without any control of such data sharing, we can only increase  
%replication for global datasets to enhance the availability,
%but this incurs significantly higher costs.
%Thus a key goal  in our design
%is restricting the impact of  data loss to a limited number of VMs. 
%It is not desirable that small scale of data failure affects the backup of many VMs.
%when users access snapshots in a recovery process. 
\item{\bf Packaging  data chunks as file system blocks.}
Since the file block size in the Hadoop and GFS is uniform and large with 64MB as a default setting,
the content chunk in a deduplication system is of nonuniform size with 4KB or 8KB on average.
We need to build an intermediate layer that supports large snapshot writing with append operations
and infrequent snapshot access.
Packaging that maps data chunks to file system blocks can create data dependence among VMs
since a file block can be shared even more VMs.
Thus we need to consider a minimum association of a fie system block to VMs in the packaging process.
%The distributed file system~\cite{GoogleFS,Hadoop} provides the replication support
%and we can leverage existing open-source software to implement such a layer.
%When the number of machines failed exceeds the replication degree, there could be
% some data loss.  
%Storage deduplication worsens the impact of such faults
%because  the majority of each snapshot (e.g. over 90\%) is deduplicated
%and the corresponding data chunks stored in the underlying file blocks
%are shared by many VMs. 
%The failure of one data chunk or block can impact many sharing VMs. 
\end{itemize}

@@

A deduplication scheme compares the fingerprints of the current snapshot
with its parent snapshot and also other snapshots in the entire cloud without consideration of .
%The traditional approach compares all fingerprints and
%stores one copy for all of a chunk's duplicates.
We call this as  the VM-oblivious (VO) approach.
In designing and selecting a duplication algorithm, we have considered the following options.
\begin{itemize}
\item {\bf Version-based change detection}.
VM snapshots can be  backed up  incrementally by identifying file  blocks that have
changed from the previous version of the snapshot~\cite{Clements2009,Vrable2009,TanIPDPS2011}.
Active snapshots will contain all the information needed to restore the virtual disk
and when deleting a snapshot, data referenced by other snapshots are removed.
While full signature comparison can deliver additional reduction~\cite{Guo2011,Dong,ExtremeBining},
content change detection falls into a VM-centric approach since deduplication is localizable. 
We are seeking for additional optimization to improve duplication efficiency and design the 
association of data chunks with the underlying file blocks so that
file block sharing among VMs is minimized.
\item {\bf sampled Index}.
One alternative approach to reducing the use of memory space is 
to use a sampled index with prefetching, proposed  by Guo and Efstathopoulos\cite{Guo2011}. 
An evaluation with our test data shows that using a sampled index can achieve a high
deduplication efficiency with a single machine setup.  
One key problem is that the algorithm is VM oblivious and not easy to be adopted for VM-centric.
Another problem is  to design a distributed version 
in deduplicating large bodies of data. 
%The algorithm stores the sampled index in memory, which is required for high
%throughput to avoid the disk-bottleneck problem - since the index itself has no
%locality information, lookups are essentially random, so the index must be
%somehow optimized to prevent excessive disk seeking (the sampling algorithm does
%this by storing the index in memory). 
To use a distributed memory version of the sampled index, every deduplicate request
may access a remote machine for index lookup,  which  incurs a significant overhead.

Another possibility is to use this approach partially  in our VM centric solution by indexing
the most popular data chunks. For a small set of popular data chunks, the prefetching strategy
used in the sampled index will not work well because the spatial locality is limited among popular 
data chunks. In our test data,  on average the number of consecutive data chunks is 7, which is too small
to be effective for index sampling.

@@

With these considerations in mind, we study a 
VM-centric approach (called VC)
for a backup service co-hosted   in the existing set of machines and resource usage is friendly
to the existing applications.  
We will first discuss and analyze the integration of the VM-centric deduplication strategies with fault isolation, and then present
an architecture and implementation design with deletion support.
